---
title: "Project Proposal"
subtitle: "Revised Proposal"
author: "Probability Pandas"
date: "10/17/2021"
output: pdf_document
---

# Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
``` 

# Load Data
```{r load-data, message = F}

initial_data = read.csv("~/Probability Pandas Project/data/500_Cities__City-level_Data__GIS_Friendly_Format___2018_release.csv")

cities <- initial_data %>%
  select(StateAbbr, PlaceName, PlaceFIPS, Geolocation, ACCESS2_AdjPrev, 
         ACCESS2_Adj95CI, CANCER_AdjPrev, CANCER_Adj95CI, CHD_AdjPrev, 
         CHD_Adj95CI, CHECKUP_AdjPrev, CHECKUP_Adj95CI, COPD_AdjPrev, 
         COPD_Adj95CI, COLON_SCREEN_AdjPrev, COLON_SCREEN_Adj95CI, 
         COREM_AdjPrev, COREM_Adj95CI, COREW_AdjPrev, COREW_Adj95CI, 
         KIDNEY_AdjPrev, KIDNEY_Adj95CI, MAMMOUSE_AdjPrev, MAMMOUSE_Adj95CI, 
         PAPTEST_AdjPrev, PAPTEST_Adj95CI) %>%
  rename(state = StateAbbr, city = PlaceName, 
         healthcare_access = ACCESS2_AdjPrev, 
         healthcare_access_CI = ACCESS2_Adj95CI, cancer = CANCER_AdjPrev, 
         cancer_CI = CANCER_Adj95CI, heart_disease = CHD_AdjPrev, 
         heart_disease_CI = CHD_Adj95CI, checkup = CHECKUP_AdjPrev, 
         checkup_CI = CHECKUP_Adj95CI, chronic_lung_disease = COPD_AdjPrev,
         chronic_lung_disease_CI = COPD_Adj95CI, 
         colon_screen = COLON_SCREEN_AdjPrev, 
         colon_screen_CI = COLON_SCREEN_Adj95CI, 
         men_colorectal_cancer_screen = COREM_AdjPrev, 
         men_colorectal_cancer_screen_CI = COREM_Adj95CI, 
         women_colorectal_cancer_screen = COREW_AdjPrev, 
         women_colorectal_cancer_screen_CI = COREW_Adj95CI, 
         chronic_kidney_disease = KIDNEY_AdjPrev, 
         chronic_kidney_disease_CI = KIDNEY_Adj95CI, 
         mammogram = MAMMOUSE_AdjPrev, 
         mammogram_CI = MAMMOUSE_Adj95CI, pap_test = PAPTEST_AdjPrev, 
         pap_test_CI = PAPTEST_Adj95CI)

```

# Introduction and Data, including Research Questions

This data is from the US Centers for Disease Control and Prevention, Epidemiology and Surveillance Branch. It contains data of the 500 largest cities in the United States. The cases are each of the 500 cities and the variables include prevalence of lack of health insurance, prevalence of clinical preventive services, screening services, and routine doctor visits, and prevalence of various chronic diseases, including but not limited to cancer, coronary heart disease, obstructive pulmonary disease, and kidney disease. The data set also includes the 95% confidence interval for the various variables in the data.

What are ways in which large cities can help prevent its citizens from the risks and mortality of chronic disease? Would wider access to healthcare prevent deaths, and what policy actions might this require? Risks to mortality from chronic disease are related to the prevalence of screening, as seen in recent literature. We can then see the relationship of health insurance to screening, which will show a link from health insurance to mortality of various chronic diseases. Our hypothesis is that increased health insurance is correlated to increased screening, resulting in decreased mortality. The inverse also holds: decreased health insurance is correlated to decreased screening, and by extension, increased mortality.

The link between screening and the mortality of various chronic diseases is seen in the following examples:

“Women who participated in mammography screening had a statistically significant 41% reduction in their risk of dying of breast cancer within 10 years and a 25% reduction in the rate of advanced breast cancers” according to a July 1, 2020 article in the American Cancer Society Journal Cancer, titled “Mammography screening reduces rates of advanced and fatal breast cancers: Results in 549,091 women”. According to a study published in the peer-reviewed journal PLOS One, in an article titled “Colonoscopy reduces colorectal cancer mortality: A multicenter, long-term, colonoscopy-based cohort study,” the colorectal cancer mortality of patients who received a colonoscopy was significantly less than those who did not receive a colonoscopy. According to a study published in the journal BMC Public Health, cervical cancer mortality was shown to have declined following the implementation of a papanicolaou test screening program. “Fecal occult blood testing reduces colorectal cancer mortality by 16 percent,” according to a study published in a June 2007 article of the peer-reviewed journal American Family Physician. A 2014 study published in World Journal of Gastroenterology showed “a 28% risk reduction in overall CRC [colorectal cancer] mortality” among those who underwent sigmoidoscopy screening.

# Glimpse

```{r glimpse initial data}

glimpse(cities)

```

# Data Analysis Plan

The table shows the prevalence of lack of health care access in each state, as well as the prevalence for chronic disease screening in each state. This table is general summarization of the data given by city in the data set to the states that they cities reside in. The first plot explores the relationship between lack of health care access and at least yearly check ups with a doctor. While there doesn't seem to be any correlation between these two variables, we can see a correlation between lack of health care access and also a decrease in mammograms. This suggests that the two variables may be positively related. 

```{r summary-stats-visualization, message = FALSE, warning = FALSE}

# Prevalence of Lack of Healthcare Access and Chronic Disease Screening by State

cities %>%
  group_by(state) %>%
  summarise(state_prev = mean(healthcare_access), 
            colon_screen = mean(colon_screen), 
            men_colorectal_cancer_screen = mean(men_colorectal_cancer_screen), 
            women_colorectal_cancer_screen = mean(women_colorectal_cancer_screen), 
            mammogram_screen = mean(mammogram, na.rm = TRUE), 
            cervical_cancer_screen = mean(pap_test, na.rm = TRUE))

# Exploratory Visualizations

ggplot(cities) +
  geom_smooth(aes(x = healthcare_access, y = checkup)) +
  labs(title = "Healthcare Access and Regular Checkups", 
       x = "Lack of Health Care Access",y = "Checkup Within a Year")

ggplot(cities) +
  geom_smooth(aes(x = healthcare_access, y = mammogram)) +
  labs(title = "Healthcare Access and Mommogram Screenings", 
       x = "Lack of Health Care Access", y = "Mammogran Screening Prevalence")
  

```


We will be looking at a potential correlation between lack of healthcare access and screening access. The predictor variable we will use is the prevalence of lack of health care in the various cities. The outcome variables we will use are the prevalence of the various screening methods for chronic diseases. To answer our questions, we will use the statistical methods of modeling/visualizing and two sided t-tests. 

The results from our modeling/visualizing would support our hypothesis if they show a positive correlation between prevalence of health insurance and prevalence of the many screening methods. We would use the 2-sided t-test to observe any significant differences in the prevalence of health care with a variety of screening methods and their relation to chronic diseases. 
