---
title: "Health Insurance and Screening for Chronic Illnesses"
subtitle: "500 Cities Project"
author: "Probability Pandas"
institute: "Duke University"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
# Remove any packages from this list that you're not using
library(knitr)
library(xaringanthemer)
library(tidyverse)
library(tidyr)
library(dplyr)
library(plyr)
library(tidymodels)
library(infer)
library(readr)
library(ggplot2)
library(sf)
library(plotly)
library(tab)
library(pROC)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(kableExtra)
library(gtsummary)
library(repr)
library(gridExtra)
library(xtable)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(
  fig.retina = 3, 
  dpi = 300, 
  fig.width = 6, 
  fig.asp = 0.618, 
  out.width = "70%"
  )
```

```{r load-data, include=FALSE}
initial_data <- read_csv("../data/500_Cities.csv")
options(scipen=10000)

cities <- initial_data %>%
  select(StateAbbr, PlaceName, Population2010, PlaceFIPS, Geolocation, 
         ACCESS2_AdjPrev,CANCER_AdjPrev, CHD_AdjPrev, CHECKUP_AdjPrev,
         COPD_AdjPrev, COLON_SCREEN_AdjPrev,COREM_AdjPrev, COREW_AdjPrev, 
         KIDNEY_AdjPrev, MAMMOUSE_AdjPrev, PAPTEST_AdjPrev) %>%
  dplyr::rename(state = StateAbbr, city = PlaceName, population = Population2010,
         health_access = ACCESS2_AdjPrev, cancer = CANCER_AdjPrev,
         heart_disease = CHD_AdjPrev, checkup = CHECKUP_AdjPrev,
         chronic_lung_disease = COPD_AdjPrev,
         colon_screen = COLON_SCREEN_AdjPrev,
         men_colorectal_cancer_screen = COREM_AdjPrev,
         women_colorectal_cancer_screen = COREW_AdjPrev,
         chronic_kidney_disease = KIDNEY_AdjPrev,
         mammogram = MAMMOUSE_AdjPrev,
         pap_test = PAPTEST_AdjPrev) %>%
  mutate(large_metro = (population >= 1500000), 
         metro = (population >= 500000 & population < 1500000), 
         med_urban = (population >= 200000 & population < 500000), 
         small_urban = (population >= 50000 & population < 200000)) %>%
  mutate(west = (state %in% c("WA", "OR", "ID", "MT", "WY", "CA", "NV", "UT", 
                              "CO", "AZ", "NM")), 
         midwest = (state %in% c("ND", "SD", "NE", "KS", "MN", "IA", "MO", 
                                 "WI", "IL", "IN", "MI", "OH")),
         northeast = (state %in% c("PA", "NY", "VT", "NH", "MA", "CT", "RI", 
                                   "NJ", "ME", "DC")), 
         south = (state %in% c("OK", "TX", "AR", "LA", "MS", "AL", "TN", "KY", 
                               "WV", "VA", "MD", "DE", "NC", "SC", "GA", 
                               "FL"))) %>%
  mutate(region = ifelse(west == TRUE, "West", 
                         ifelse(midwest == TRUE, "Midwest", 
                                ifelse(northeast == TRUE, "Northeast", 
                                       ifelse(south == TRUE,  
                                              "South", NA))))) %>%
  mutate(city_size = ifelse(large_metro == TRUE, "Large Metropolitan", 
                         ifelse(metro == TRUE, "Metropolitan", 
                                ifelse(med_urban == TRUE, "Medium-Size Urban", 
                                       ifelse(small_urban == TRUE, 
                                              "Small-Size Urban", NA))))) %>%
  na.omit(city_size) %>%
  na.omit(region) %>%
  mutate(checkup_n = (population*(checkup/100)), 
         colon_n_screened = (population*(colon_screen/100)), 
         m_colorectal_n_screened = (population*(men_colorectal_cancer_screen/100)),
         w_colorectal_n_screened = 
           (population*(women_colorectal_cancer_screen/100)),
         mammogram_n_screened = (population*(mammogram/100)),
         pap_n_screened = (population*(pap_test/100))) %>%
  mutate(no_checkup_n = (population - checkup_n), 
         no_colon_n_screened = (population - colon_n_screened), 
         no_m_colorectal_n_screened = (population - m_colorectal_n_screened),
         no_w_colorectal_n_screened = 
           (population - w_colorectal_n_screened),
         no_mammogram_n_screened = (population - mammogram_n_screened),
         no_pap_n_screened = (population - pap_n_screened))

#this is the training data set called model_cities
set.seed(100)

split <- initial_split(cities, prop = 3/4, strata = region)

model_cities <- training(split)
cities_test <- testing(split)

#for introductory visualizations
west_cities <- cities %>%
  filter(west)

midwest_cities <- cities %>%
  filter(midwest)

northeast_cities <- cities %>%
  filter(northeast)

south_cities <- cities %>%
  filter(south)


#wrangling for visualizations
cities2 <- cities %>%
  filter( region != "NA") %>%
  mutate(logcheckup_n = log(checkup_n))
```

```{r style-slides, echo = FALSE}
style_xaringan(
  title_slide_background_image = "img/watercolour_sys02_img34_teacup-ocean.jpg"
)
```

class: inverse, middle, center

# 500 Cities Project Database

---

## Background: Screening for Chronic Disease

We found several studies revealing that obtaining a screening test for breast cancer, colorectal cancer, colon cancer, and cervical cancer would reduce a patient's risk of mortality by his or her chronic disease. For example, the American Cancer society conducted a study to observe the effects of mammogram screenings on the prevalence of advanced and fatal breast cancers. The study found that women who received breast cancer screening had a statistically significant 41% reduction in their risk of having a fatal breast cancer case and a 25% reduction in the rates of having advanced breast cancer. This study suggests that receiving a screening can decrease the likelihood of having a fatal or worsened case of breast cancer.

- The goal of this report is to analyze how lack of access to health insurance affects people’s access to chronic disease screening. 

---

## Research Question

How does lack of access to health insurance correlate to the prevalence of screening for various chronic diseases?

---

## Hypothesis

Our hypothesis is that decreased lack of access to health insurance prevalence is correlated to increased screening for chronic diseases in various cities. Inversely, we also hypothesize that increased lack of access to health insurance prevalence is correlated to decreased screening for chronic diseases in various cities.

- Null Hypothesis: The lack of health insurance has no effect on the number of chronic disease screenings in various cities.

- Alternate Hypothesis:  The lack of health insurance has an effect on the number of chronic disease screenings in various cities.

---

class: inverse, middle, center

# Data Wrangling

---

## New Variables

- Regions: West, Midwest, North, and South
- City Sizes
   Large Metropolitan: population of 1.5 million or more, 
   Metropolitan: population between 500,000-1.5 million,          
   Medium-Size Urban: population was between 200,000-500,000,  
   Small Urban: population between 50,000-200,000
- Number of Individuals Getting Specific Screening

---

class: inverse, middle, center

# Exploratory Data Analysis

---

## Correlation Between Checkups and Screening for Chronic Diseases

```{r correlation-table, warning=FALSE, message=FALSE, echo=FALSE}
correlation1 <- cor.test(cities$checkup_n, cities$pap_n_screened,
                         method = "pearson")
tidy.cor1 <- tidy(correlation1) %>%
  rename("Correlation Coefficient Estimate" = estimate, "P-Value" = p.value, 
         "CI Low" = conf.low, "CI High" = conf.high)
tidy.cor1.final <- tidy.cor1[,-c(2,4,7,8)]
kable(tidy.cor1.final, caption = "Pearson's Correlation Test: Pap Tests and Checkups") %>%
    row_spec(0, bold = T) %>%
  kable_styling(latex_options = "HOLD_position")

correlation2 <- cor.test(cities$checkup_n, cities$colon_n_screened,
                         method = "pearson")
tidy.cor2 <- tidy(correlation2) %>%
  rename("Correlation Coefficient Estimate" = estimate, "P-Value" = p.value, 
         "CI Low" = conf.low, "CI High" = conf.high)
tidy.cor2.final <- tidy.cor2[,-c(2,4,7,8)]
kable(tidy.cor2.final, caption = "Pearson's Correlation Test: Colon Screenings and Checkups") %>%
    row_spec(0, bold = T) %>%
  kable_styling(latex_options = "HOLD_position")

correlation3 <- cor.test(cities$checkup_n, cities$mammogram_n_screened,
                         method = "pearson")
tidy.cor3 <- tidy(correlation3) %>%
    rename("Correlation Coefficient Estimate" = estimate, "P-Value" = p.value, 
         "CI Low" = conf.low, "CI High" = conf.high)
tidy.cor3.final <- tidy.cor3[,-c(2,4,7,8)]
kable(tidy.cor3.final, caption = "Pearson's Correlation Test: Mammograms and Checkups") %>%
    row_spec(0, bold = T) %>%
  kable_styling(latex_options = "HOLD_position")
```

---

## Checkups by Region
```{r checkups-by-region1, echo = FALSE, warning = FALSE, message = FALSE}
ggplot(data = cities2,
       aes(x = region, y = logcheckup_n,
           color = region)) +
  geom_boxplot() +
  labs(x = "Region",
       y = "Checkups per capita (log scale)",
       title = "Checkups by Region") +
  theme(axis.text.x = element_text(angle = 45,hjust=1), legend.position = "none")
```

---

## Checkups by Region
```{r anova, echo = FALSE, warning = FALSE, message = FALSE}
anova1 <- aov(logcheckup_n ~ region, data = cities2)
tidy.anova1 <- tidy(anova1) %>%
      rename("Term" = term, "Statistic" = statistic, "P-Value" = p.value)
tidy.anova1.final <- tidy.anova1[-c(2),-c(2,3,4)]
kable(tidy.anova1.final, caption = "ANOVA: Checkups per capita (log scale), by Region") %>%
      row_spec(0, bold = T) %>%
  kable_styling(latex_options = "HOLD_position")
```

---

## Checkups by City Size
```{r Simple-Marginal-Relationship4, warning=FALSE, message=FALSE, echo=FALSE}
#(4) boxplots of checkup (n and log) by city size
ggplot(data = cities2,
       aes(x = city_size, y = logcheckup_n,
           color = city_size)) +
  geom_boxplot() +
  labs(x = "City Size",
       y = "Checkups per capita (log scale)",
       title = "Checkups by City Size") +
  theme(axis.text.x = element_text(angle = 45,hjust=1), legend.position = "none")
```

---

## Checkups by City Size
```{r anova2, warning=FALSE, message=FALSE, echo=FALSE}
#(6) anova shows means of checkup (n and log) are different among the city sizes
anova2 <- aov(logcheckup_n ~ city_size, data = cities2)
tidy.anova2 <- tidy(anova2) %>%
        rename("Term" = term, "Statistic" = statistic, "P-Value" = p.value)
tidy.anova2.final <- tidy.anova2[,-c(2,3,4)]
kable(tidy.anova2.final, caption = "ANOVA: Checkups per capita (log scale), by Region") %>%
      row_spec(0, bold = T) %>%
  kable_styling(latex_options = "HOLD_position")

```

---
class: inverse, middle, center

# Modeling

---

## Multiple Logistic Regression Model
```{r Multiple-Logistic-Regression-Model, echo=FALSE, warning=FALSE, message=FALSE}

#model 1: withouth interaction
checkup_fit <- glm(cbind(checkup_n, no_checkup_n) ~ region + 
                     city_size + health_access, 
                   family = binomial, data = model_cities)
checkup_fit_df1 <- tidy(checkup_fit, exponentiate = TRUE, intercept = TRUE) %>%
  rename(Variable = term, Estimate = estimate, "P-Value" = p.value) %>%
  mutate(Variable = c("Intercept", "Northeast", "South", "West",
                                  "Medium-Size Urban", "Metropolitan",
                                  "Small-Size Urban", "HI"))
checkup_fit_df1_f <- checkup_fit_df1[, -c(3, 4)]

#model 2: with interaction
checkup_fit2 <- glm(cbind(checkup_n, no_checkup_n) ~ region +
                     city_size + health_access + region*city_size +
                      region*health_access + city_size*health_access,
                   family = binomial, data = model_cities) 

checkup_fit_df2 <- tidy(checkup_fit2, exponentiate = TRUE, intercept = TRUE) %>%
  rename(Variable = term, Estimate = estimate, "P-Value" = p.value) %>%
  mutate(Variable = c("Intercept", "Northeast", "South", "West",
                      "Medium-Size Urban", "Metropolitan", "Small-Size Urban", 
                      "HI", "Northeast by Medium-Size Urban", 
                      "South by Medium-Size Urban", 
                      "West by Medium-Size Urban", "Northeast by Metropolitan", 
                      "South by Metropolitan", "West by Metropolitan", 
                      "Northeast by Small-Size Urban", 
                      "South by Small-Size Urban", "West by Small-Size Urban", 
                      "Northeast by HI", "South by HI", "West by HI", 
                      "Medium-Size Urban by HI", "Metropolitan by HI", 
                      "Small-Size Urban by HI"))
checkup_fit_df2_f <- checkup_fit_df2[c(2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 
                                       12, 13, 14), -c(3, 4)]

```

We created a Logistic Regression Model because we wanted to model the binary outcome of whether or not individuals in a certain city were getting checkups or not. 

Model 1: Without Interaction
$$log(\frac{\pi_{i}}{\pi-1})=\beta_{0} + \beta_{1}*Region + \beta_{2}*CitySize + \beta_{3}HealthInsurance$$
Model 2: With Interaction
$$log(\frac{\pi_{i}}{\pi-1})=\beta_{0} + \beta_{1}*Region + \beta_{2}*CitySize + \beta_{3}HealthInsurance + \beta_{1}*Region*CitySize +$$

$$\beta_{2}*Region*HealthInsurance + \beta_{3}CitySize*HealthInsurance$$

---

class: inverse, middle, center

# Model Validation

```{r model-validation, echo=FALSE, warning=FALSE, message=FALSE}
# predictions from no interaction model 
test_prob1 = predict(checkup_fit, newdata = cities_test, type = "response")

cities_test1 <-
  merge(cities_test, test_prob1, by = "row.names", all.x = TRUE)

# prediction from model with interaction
test_prob2 = predict(checkup_fit2, newdata = cities_test, type = "response")

cities_test2 <-
  merge(cities_test, test_prob2, by = "row.names", all.x = TRUE)

```

---

## RMSE Against Training Set

```{r rmse-training, echo=FALSE, warning=FALSE, message=FALSE}
#rmse against model/training data set

#model 1
test_prob_model = predict(checkup_fit, newdata = model_cities, type = "response")
model_cities1 <- merge(model_cities, test_prob_model, by = "row.names", all.x = TRUE)

rmse1 <- rmse(model_cities1, truth = (checkup/100), estimate = test_prob_model) %>%
  rename("RMSE" = .metric, "Estimate" = .estimate)
rmse1_f <- rmse1[, c(1, 3)]

#model 2
test_prob_model2 = predict(checkup_fit2, newdata = model_cities, type = "response")
model_cities2 <- merge(model_cities, test_prob_model2, by = "row.names", all.x = TRUE)

rmse2 <- rmse(model_cities2, truth = (checkup/100), estimate = test_prob_model)%>%
  rename("RMSE" = .metric, "Estimate" = .estimate)
rmse2_f <- rmse2[, c(1, 3)]

```

.pull-left[
```{r rmse-table1, echo=FALSE, warning=FALSE, message=FALSE}

rmse1_f %>%
  kable(caption = "RMSE of Model 1 Against Traning Dataset")%>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "HOLD_position")

```
]

.pull-right[
```{r rmse-table2, echo=FALSE, warning=FALSE, message=FALSE}

rmse2_f %>%
  kable(caption = "RMSE of Model 2 Against Traning Dataset") %>%
  kable_classic(full_width = F, html_font = "Cambria")%>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "HOLD_position")

```
]

---

## RMSE Against Test Set

```{r rmse-test, echo=FALSE, warning=FALSE, message=FALSE}
#calculate root mean standard error of both models against test data set

#model 1
rmse3 <- rmse(cities_test1, truth = (checkup/100), estimate = test_prob1) %>%
  rename("RMSE" = .metric, "Estimate" = .estimate)
rmse3_f <- rmse3[, c(1, 3)]

#model 2
rmse4 <- rmse(cities_test2, truth = (checkup/100), estimate = test_prob2) %>%
  rename("RMSE" = .metric, "Estimate" = .estimate)
rmse4_f <- rmse4[, c(1, 3)]
```

.pull-left[
```{r rmse-table3, echo=FALSE, warning=FALSE, message=FALSE}

rmse3_f %>%
  kable(caption = "RMSE of Model 1 Against Test Dataset")%>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "HOLD_position")

```
]

.pull-right[
```{r rmse-table4, echo=FALSE, warning=FALSE, message=FALSE}

rmse4_f %>%
  kable(caption = "RMSE of Model 2 Against Test Dataset")%>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "HOLD_position")

```
]
---

## Results

```{r model1-table-again, echo=FALSE}

checkup_fit_df1_f %>%
  kable(caption = "Model 1: Without Interaction") %>%
  kable_classic(full_width = F, html_font = "Cambria") %>%
  row_spec(0, bold = T) %>%
  kable_styling(latex_options = "HOLD_position")

```

---

## Discussion

- Impact of Different Regions on Predicted Checkups
- Impact of Different City Sizes on Predicted Checkups
- Impact of HI on Predicted Checkups

---

## Limitations

- Unequal number of cities for each city size
- Data set only includes 500 largest cities in the US
- Other significant predictors that are not included in this particular dataset

---

## Potential Implications

- Suggestions for further research: Why are cities in the West less likely to have a high prevalence of checkups?
- Results of the model as it pertains to prevalence of lack of health insurance: Increased prevalence of health insurance is correlated to greater amount of checkups per capita, and therefore to increased prevalence of screening for various diseases.
- Mortality: Studies show link from increased screening to decreased mortality.
- Lowered mortality rates may be achieved with greater prevalence of health insurance.

---
